Analysing the effect of land-use type on the proportion of scouts

### 2022-01-27 analysis/fit_models_durations.Rmd
changed analysis to fit models to waggle run duration

This document contains code to fit the waggle dance model to each of our 20 different sites. The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here. Each site has its own section, with reused code. The final sections create map plots and summary statistic plots.

@@

@@

# subset data for target site
data <- get_data() %>%
  filter(site == target_site)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 10, -5.5, -2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 4.5, -6.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 12, -8., -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 5.5, -6.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 2.5, -5, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 2.7, -8., -2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 5, -6.5, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 6, -6.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 4, -9, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 5, -6.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 3.5, -7, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 4.5, -7.2, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 3., -5.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 3.5, -6.5, -2.5)

@@

# subset data for target site
data <- get_data() %>%
  filter(site == target_site) %>%
  filter(foraging_distance < 9) # remove outlier distance

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 4., -7.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 5.5, -6.5, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 2.5, -6, -2.)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 5., -8, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.5, 2.4, -8, -2)

>>>>

REMEBER: Once we know how we are doing the conversion from duration to distance, or what we are doing with it, reset this branch to the remote one and check everthing is as was. Then make the edits in the dev branch, merge the relevent bits in properly and then redo the analysis, remaking the figures for the code. You MUST reset this to remote to prevent overwiting graphs and what not.

This document contains code to fit the waggle dance model to each of our 20 different sites. The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things. Each site has its own section, with reused code. The final sections create map plots and summary statistic plots.

@@

alldata <- data %>%
  select(date, site, duration.seconds) %>%
  rename(foraging_distance = duration.seconds)

head(alldata)

@@

# subset data for target site
data <- alldata %>%
  filter(site == target_site)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 8, -5.5, -1.8)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.2, 3.5, -6.5, -2.)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.2, 9.5, -8., -2.2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4, -6.5, -2.2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 1.8, -6, -1.2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 2.1, -8., -2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4, -6.5, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4, -6.5, -2.2)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 3, -9, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4.2, -6.5, -2.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 3., -7, -2.1)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4., -7.2, -2.1)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 2., -5.5, -2.1)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 2.5, -6.5, -2.1)

@@

# subset data for target site
data <- alldata %>%
  filter(site == target_site) %>%
  filter(foraging_distance < 6) # remove outlier distance

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 3., -7.5, -2.1)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4, -6.5, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 2., -6, -1.5)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 4., -8, -2.)

@@

# set coordinates for histogram subplot
subplot_coords <- c(0.1, 1.8, -8, -2)

### 2022-01-27 analysis/fit_models_newdist.Rmd
changed analysis to fit models to waggle run duration

@@

# subset data for target site
data <- get_data() %>%
  filter(site == target_site) %>%
  filter(foraging_distance < 9) # remove outlier distance

>>>>

alldata <- data %>%
  select(date, site, newforaging_distance) %>%
  rename(foraging_distance = newforaging_distance)

@@

# subset data for target site
data <- alldata %>%
  filter(site == target_site) %>%
  filter(foraging_distance < 9) # remove outlier distance

### 2022-01-27 analysis/landuse_effect_prop_scouts.Rmd
changed analysis to fit models to waggle run duration

# select sites to draw attention to, those sitting on the trend
ar_marked <- ar_sortedpropscouts %>%
  filter(site %in% c("MEL", "SOM", "BEL"))

@@

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  ylim(-0.6, 0.6)

@@

@@

```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}

@@

# make fit map with agrirural landuses
urban_plotdata$site <- urban_lu$site
urban_plotdata$prop_scouts <- urban_lu$p
u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)
u_marked <- u_sortedpropscouts %>%
  filter(site %in% c("HOR", "SWP", "CAD"))

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_u <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(urban_luconversions, by = c("descriptio" = "code"))
  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "continuous central" = "#EE6677", "dense residential" = "#AA3377",
        "sparse residential" = "#CCBB44", "amenity grassland" = "#228833",
        "water" = "#4477AA", "parks allotments cemeteries" = "#BBBBBB",
        "Railway" = "#BBBBBB", "woodland" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.55,
    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#EE6677", "dense.residential" = "#AA3377",
      "sparse.residential" = "#CCBB44", "amenity.grassland" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  scale_y_continuous(breaks = seq(-1, 1, by = 0.4), limits = c(-1, 1))

>>>>

print(ar_sortedpropscouts)

# select sites to draw attention to, those sitting on the trend
ar_marked <- ar_sortedpropscouts %>%
  filter(site %in% c("MEL", "ROT", "BEL"))

@@

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, -0.05), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  ylim(-0.6, 0.6)

@@

ggsave(
  plot = comb,
  filename = "results/figures/lu_prop_scouts_agrirural.svg",
  width = 183,
  height = 183,
  units = "mm",
  dpi = 300
)

@@

```{r, urban_landuse_maps, message = FALSE, warning = FALSE}

@@

# make fit map with agrirural landuses
urban_plotdata$site <- urban_lu$site
urban_plotdata$prop_scouts <- urban_lu$p
u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)

print(u_sortedpropscouts)

u_marked <- u_sortedpropscouts %>%
  filter(site %in% c("HHS", "MPA", "CAD"))

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_u <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(urban_luconversions, by = c("descriptio" = "code"))
  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "continuous central" = "#BBBBBB", "dense residential" = "#BBBBBB",
        "sparse residential" = "#EE6677", "amenity grassland" = "#BBBBBB",
        "water" = "#BBBBBB", "parks allotments cemeteries" = "#228833",
        "Railway" = "#CCBB44", "woodland" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.3), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  theme(text = element_text(size = 10)) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -urban_loadings$cutoff,
      ymax = urban_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#BBBBBB", other = "#BBBBBB"
    )
  ) +
  scale_y_continuous(breaks = seq(-1, 1, by = 0.4), limits = c(-1, 1))
